### 2011-02-10 frame.tex
First commit

\documentclass{elsart}
\usepackage[pdftex]{hyperref}
\usepackage{amstext,amsmath,amsfonts,amssymb}
\usepackage{bm}
\usepackage{lscape}
\usepackage{graphicx}
\usepackage{vruler}
\usepackage{color}
\usepackage{subfigure}
\usepackage{mathrsfs}
\usepackage{eufrak}
\usepackage{float}
\usepackage{rotating}
\textwidth 17.0cm
\oddsidemargin 0.0cm
\evensidemargin 0.0cm
\topmargin 0.5cm
% ===========================================================================

\begin{document}
\newcommand{\blue}[1]{{\color{blue}{#1}\color{black}}}
\newcommand{\red}[1]{{\color{red}{#1}\color{black}}}

\newcommand{\green}[1]{{\color{green}{#1}\color{black}}}
\newcommand{\yellow}[1]{{\color{yellow}{#1}\color{black}}}
\newcommand{\darkblue}[1]{{\color{darkblue}{#1}\color{black}}}
\newcommand{\darkgreen}[1]{{\color{darkgreen}{#1}\color{black}}}
\newcommand{\darkred}[1]{{\color{darkred}{#1}\color{black}}}
\newcommand{\lightred}[1]{{\color{lightred}{#1}\color{black}}}
\newcommand{\pink}[1]{{\color{pink}{#1}\color{black}}}
\newcommand{\gray}[1]{{\color{gray}{#1}\color{black}}}
\newcommand{\black}[1]{{\color{black}{#1}\color{black}}}
\newcommand{\magenta}[1]{\color{magenta}{#1}\color{black}}
\newcommand{\darkmag}[1]{\color{darkmag}{#1}\color{black}}

\def\CalT{\mathcal{T}}
\def\rmBH{\mathrm{BH}}
\def\rmDVCS{\mathrm{DVCS}}
\def\CalI{\mathcal{I}}
\def\rmunp{\mathrm{unp}}
\def\rmTP{\mathrm{TP}}
\def\rmRe{\mathrm{Re}}
\def\rmIm{\mathrm{Im}}

\def\xbj{x_{B}}
\def\qsr{Q^{2}}

\def\gpdH{\emph{H}}
\def\gpdE{\emph{E}}
\def\gpdHtil{\widetilde{H}}
\def\gpdEtil{\widetilde{E}}

\def\CalH{\mathcal{H}}
\def\CalE{\mathcal{E}}
\def\CalHtil{\widetilde{\mathcal{H}}}
\def\CalEtil{\widetilde{\mathcal{E}}}

\def\CalN{\mathcal{N}}
\def\CUU{\sigma_{U}}
\def\Lumi{\mathcal{L}\,}
\def\xx{\bm{x}}
\def\tt{\bm{\eta}}
\def\AC{A_C}
\def\AUT{A_{UT}}
\def\AUTDVCS{A_{UT,\rmDVCS}}
\def\AUTI{A_{UT,\CalI}}
\def\desy{{\sc desy}}

 \begin{frontmatter}

\title{
	Beam-Helicity and Charge Asymmetries \\
        in the Bethe-Heitler and DVCS processes on \\
        an Unpolarised Proton Target
  	\underline{Paper Tag: PDVCS (v0.5)}
        \underline{Drafting Committee: DC90 (\today)}
}

\collab{HERMES Collaboration}
\begin{abstract}
Beam\blue{-}helicity and charge asymmetries in the hard exclusive leptoproduction of
real photons from an unpolarised proton target \blue{by} a 27.6\,GeV lepton beam 
are extracted from the HERMES unpolarised data set of 2006-2007 using a
missing-mass event selection technique. The beam-charge asymmetry amplitudes extracted from the new data set are statistically more precise than those extracted from an earlier data set of 1996-2005 previously analysed in the same manner by the collaboration. \blue{Results from both data sets are found to be} compatible with each other. Combined results from these data sets are extracted and are the statistically most precise asymmetry amplitude measurements available in the HERMES kinematic region.
\end{abstract}

\end{frontmatter}
% 
\setvruler[15pt][1][1][3][1][30pt][30pt][-42pt]
\include{theory}
\include{experiment_and_data}
\include{background_and_systematic}
\include{results}
\include{summary}
\include{append}
% this is a test comment to test the pull push mechanism
\bibliographystyle{ieeetr}
\bibliography{jburnsref}
\end{document}