### 2011-03-29 frame.tex
Added the JHEP style

\documentclass[11pt,a4paper]{article}
\usepackage{jheppub}
\pdfoutput=1
\toccontinuoustrue
\usepackage{lscape}
\usepackage{vruler}
\usepackage{subfigure}
\usepackage{appendix}

\newcommand{\blue}[1]{{\color{blue}{#1}\color{black}}}
\def\desy{{\sc desy}}

\begin{document}
\title{Beam-Helicity and Charge Asymmetries \\
        in the Bethe-Heitler and DVCS processes on \\
        an Unpolarised Proton Target
  	\underline{Paper Tag: PDVCS (v0.9)}
        \underline{Drafting Committee: DC90 (\today)}
}

% List of Institute Addresses 
\collaboration{The HERMES Collaboration}
\collaborationImg{\includegraphics{hermeslogo.pdf}}

\author[12]{A.~Airapetian}
\author[26]{N.~Akopov}
\author[5]{Z.~Akopov}
\author[6]{E.C.~Aschenauer}
\author[25]{W.~Augustyniak}
\author[26]{R.~Avakian}
\author[26]{A.~Avetissian}
\author[5]{E.~Avetisyan}
\author[15]{B.~Ball}
\author[18]{S.~Belostotski}
\author[10]{N.~Bianchi}
\author[17,24]{H.P.~Blok}
\author[6]{H.~B\"ottcher}
\author[5]{A.~Borissov}
\author[13]{J.~Bowles}
\author[12]{I.~Brodski}
\author[19]{V.~Bryzgalov}
\author[13]{J.~Burns}
\author[9]{M.~Capiluppi}
\author[10]{G.P.~Capitani}
\author[21]{E.~Cisbani}
\author[9]{G.~Ciullo}
\author[9]{M.~Contalbrigo}
\author[9]{P.F.~Dalpiaz}
\author[5]{W.~Deconinck}
\author[2]{R.~De~Leo}
\author[5,22]{L.~De~Nardo}
\author[10]{E.~De~Sanctis}
\author[8]{M.~Diefenthaler}
\author[10]{P.~Di~Nezza}
\author[17]{J.~Dreschler}
\author[12]{M.~D\"uren}
\author[12]{M.~Ehrenfried}
\author[26]{G.~Elbakian}
\author[4]{F.~Ellinghaus}
\author[6]{R.~Fabbri}
\author[10]{A.~Fantoni}
\author[22]{L.~Felawka}
\author[21]{S.~Frullani}
\author[6]{D.~Gabbert}
\author[19]{G.~Gapienko}
\author[19]{V.~Gapienko}
\author[21]{F.~Garibaldi}
\author[5,18,22]{G.~Gavrilov}
\author[26]{V.~Gharibyan}
\author[9]{F.~Giordano}
\author[15]{S.~Gliske}
\author[5]{M.~Golembiovskaya}
\author[10]{C.~Hadjidakis}
\author[5]{M.~Hartig}
\author[10]{D.~Hasch}
\author[23]{T.~Hasegawa}
\author[13]{G.~Hill}
\author[6]{A.~Hillenbrand}
\author[13]{M.~Hoek}
\author[5]{Y.~Holler}
\author[6]{I.~Hristova}
\author[23]{Y.~Imazu}
\author[19]{A.~Ivanilov}
\author[18]{A.~Izotov}
\author[1]{H.E.~Jackson}
\author[18]{A.~Jgoun}
\author[11]{H.S.~Jo}
\author[14]{S.~Joosten}
\author[13]{R.~Kaiser}
\author[26]{G.~Karyan}
\author[12]{T.~Keri}
\author[4]{E.~Kinney}
\author[18]{A.~Kisselev}
\author[23]{N.~Kobayashi}
\author[19]{V.~Korotkov}
\author[16]{V.~Kozlov}
\author[18]{P.~Kravchenko}
\author[7]{V.G.~Krivokhijine}
\author[2]{L.~Lagamba}
\author[14]{R.~Lamb}
\author[17]{L.~Lapik\'as}
\author[13]{I.~Lehmann}
\author[9]{P.~Lenisa}
\author[14]{L.A.~Linden-Levy}
\author[11]{A.~L\'opez~Ruiz}
\author[15]{W.~Lorenzon}
\author[5]{X.-G.~Lu}
\author[23]{X.-R.~Lu}
\author[3]{B.-Q.~Ma}
\author[13]{D.~Mahon}
\author[14]{N.C.R.~Makins}
\author[18]{S.I.~Manaenkov}
\author[21]{L.~Manfr\'e}
\author[3]{Y.~Mao}
\author[25]{B.~Marianski}
\author[4]{A.~Martinez de la Ossa}
\author[26]{H.~Marukyan}
\author[22]{C.A.~Miller}
\author[23]{Y.~Miyachi}
\author[26]{A.~Movsisyan}
\author[10]{V.~Muccifora}
\author[13]{M.~Murray}
\author[8]{A.~Mussgiller}
\author[2]{E.~Nappi}
\author[18]{Y.~Naryshkin}
\author[8]{A.~Nass}
\author[6]{M.~Negodaev}
\author[6]{W.-D.~Nowak}
\author[9]{L.L.~Pappalardo}
\author[12]{R.~Perez-Benito}
\author[8]{N.~Pickert}
\author[8]{M.~Raithel}
\author[1]{P.E.~Reimer}
\author[10]{A.R.~Reolon}
\author[6]{C.~Riedl}
\author[8]{K.~Rith}
\author[13]{G.~Rosner}
\author[5]{A.~Rostomyan}
\author[14]{J.~Rubin}
\author[11]{D.~Ryckbosch}
\author[19]{Y.~Salomatin}
\author[20]{F.~Sanftl}
\author[20]{A.~Sch\"afer}
\author[11]{G.~Schnell}
\author[5]{K.P.~Sch\"uler}
\author[13]{B.~Seitz}
\author[23]{T.-A.~Shibata}
\author[7]{V.~Shutov}
\author[9]{M.~Stancari}
\author[9]{M.~Statera}
\author[8]{E.~Steffens}
\author[17]{J.J.M.~Steijger}
\author[12]{H.~Stenzel}
\author[6]{J.~Stewart}
\author[8]{F.~Stinzing}
\author[26]{S.~Taroian}
\author[16]{A.~Terkulov}
\author[25]{A.~Trzcinski}
\author[11]{M.~Tytgat}
\author[11]{A.~Vandenbroucke}
\author[17]{P.B.~van~der~Nat}
\author[11]{Y.~Van~Haarlem}
\author[11]{C.~Van~Hulse}
\author[5]{M.~Varanda}
\author[18]{D.~Veretennikov}
\author[18]{V.~Vikhrov}
\author[2]{I.~Vilardi}
\author[8]{C.~Vogel}
\author[3]{S.~Wang}
\author[8]{S.~Yaschenko}
\author[3]{H.~Ye}
\author[5]{Z.~Ye}
\author[22]{S.~Yen}
\author[12]{W.~Yu}
\author[8]{D.~Zeiler}
\author[5]{B.~Zihlmann}
\author[25]{P.~Zupranski}

\affiliation[1]{Physics Division, Argonne National Laboratory, Argonne, Illinois 60439-4843, USA}
\affiliation[2]{Istituto Nazionale di Fisica Nucleare, Sezione di Bari, 70124 Bari, Italy}
\affiliation[3]{School of Physics, Peking University, Beijing 100871, China}
\affiliation[4]{Nuclear Physics Laboratory, University of Colorado, Boulder, Colorado 80309-0390, USA}
\affiliation[5]{DESY, 22603 Hamburg, Germany}
\affiliation[6]{DESY, 15738 Zeuthen, Germany}
\affiliation[7]{Joint Institute for Nuclear Research, 141980 Dubna, Russia}
\affiliation[8]{Physikalisches Institut, Universit\"at Erlangen-N\"urnberg, 91058 Erlangen, Germany}
\affiliation[9]{Istituto Nazionale di Fisica Nucleare, Sezione di Ferrara and Dipartimento di Fisica, Universit\`a di Ferrara, 44100 Ferrara, Italy}
\affiliation[10]{Istituto Nazionale di Fisica Nucleare, Laboratori Nazionali di Frascati, 00044 Frascati, Italy}
\affiliation[11]{Department of Subatomic and Radiation Physics, University of Gent, 9000 Gent, Belgium}
\affiliation[12]{Physikalisches Institut, Universit\"at Gie{\ss}en, 35392 Gie{\ss}en, Germany}
\affiliation[13]{SUPA, School of Physics and Astronomy, University of Glasgow, Glasgow G12 8QQ, United Kingdom}
\affiliation[14]{Department of Physics, University of Illinois, Urbana, Illinois 61801-3080, USA}
\affiliation[15]{Randall Laboratory of Physics, University of Michigan, Ann Arbor, Michigan 48109-1040, USA }
\affiliation[16]{Lebedev Physical Institute, 117924 Moscow, Russia}
\affiliation[17]{National Institute for Subatomic Physics (Nikhef), 1009 DB Amsterdam, The Netherlands}
\affiliation[18]{Petersburg Nuclear Physics Institute, Gatchina, Leningrad region, 188300 Russia}
\affiliation[19]{Institute for High Energy Physics, Protvino, Moscow region, 142281 Russia}
\affiliation[20]{Institut f\"ur Theoretische Physik, Universit\"at Regensburg, 93040 Regensburg, Germany}
\affiliation[21]{Istituto Nazionale di Fisica Nucleare, Sezione Roma 1, Gruppo Sanit\`a and Physics Laboratory, Istituto Superiore di Sanit\`a, 00161 Roma, Italy}
\affiliation[22]{TRIUMF, Vancouver, British Columbia V6T 2A3, Canada}
\affiliation[23]{Department of Physics, Tokyo Institute of Technology, Tokyo 152, Japan}
\affiliation[24]{Department of Physics and Astronomy, Vrije Universiteit, 1081 HV Amsterdam, The Netherlands}
\affiliation[25]{Andrzej Soltan Institute for Nuclear Studies, 00-689 Warsaw, Poland}
\affiliation[26]{Yerevan Physics Institute, 375036 Yerevan, Armenia}

\abstract{
Beam-helicity and charge asymmetries in the hard exclusive leptoproduction of real photons from an unpolarised proton target by a 27.6\,GeV lepton beam 
are extracted from the \blue{H{\sc ermes}} data set of 2006-2007 using a missing-mass event selection technique. The beam-charge asymmetry amplitudes extracted from the new data set are statistically more precise than those extracted from \blue{the} earlier data set of 1996-2005 previously analysed in the same manner by the collaboration. Results from \blue{the two} data sets are found to be compatible with each other. Combined results from these data sets are extracted and \blue{represent} the statistically most precise asymmetry amplitude measurements available in the \blue{H{\sc ermes}} kinematic region.
}
\keywords{lepton-nucleon scattering}

\setvruler[15pt][1][1][3][1][30pt][30pt][-42pt]
\maketitle

\include{theory}
\include{experiment_and_data}
\include{background_and_systematic}
\include{results}
\include{summary}
\include{append}
\bibliographystyle{natbib}
\bibliography{jburnsref}
\end{document}