### 2011-11-21 rec-dvcsP0607.tex

% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showpacs' option to make PACS codes appear
%  Add 'showkeys' option to make keywords appear

\documentclass[aps,prl,twocolumn,superscriptaddress]{revtex4}
%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4}
%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}

% Enable use of graphics

\usepackage{graphicx}

% You should use BibTeX and apsrev.bst for references
% Choosing a journal automatically selects the correct APS
% BibTeX style file (bst file), so only uncomment the line
% below if necessary.

%\bibliographystyle{apsrev}

% ===========================================================================

\begin{document}

\title{
  Beam Helicity and Charge Asymmetries in the Bethe-Heitler and Deeply-Virtual Compton Scattering Processes on an Unpol Proton Target \\
  \underline{Paper Tag: dvcsP0607}
  \hspace*{1cm}\underline{Drafting Committee: 90}
}

% List of Institute Addresses

\def\groupargonne{\affiliation{Physics Division, Argonne National Laboratory, Argonne, Illinois 60439-4843, USA}}
\def\groupbari{\affiliation{Istituto Nazionale di Fisica Nucleare, Sezione di Bari, 70124 Bari, Italy}}
\def\groupbeijing{\affiliation{School of Physics, Peking University, Beijing 100871, China}}
\def\groupcolorado{\affiliation{Nuclear Physics Laboratory, University of Colorado, Boulder, Colorado 80309-0390, USA}}
\def\groupdesy{\affiliation{DESY, 22603 Hamburg, Germany}}
\def\groupzeuthen{\affiliation{DESY, 15738 Zeuthen, Germany}}
\def\groupdubna{\affiliation{Joint Institute for Nuclear Research, 141980 Dubna, Russia}}
\def\grouperlangen{\affiliation{Physikalisches Institut, Universit\"at Erlangen-N\"urnberg, 91058 Erlangen, Germany}}
\def\groupferrara{\affiliation{Istituto Nazionale di Fisica Nucleare, Sezione di Ferrara and Dipartimento di Fisica, Universit\`a di Ferrara, 44100 Ferrara, Italy}}
\def\groupfrascati{\affiliation{Istituto Nazionale di Fisica Nucleare, Laboratori Nazionali di Frascati, 00044 Frascati, Italy}}
\def\groupgent{\affiliation{Department of Subatomic and Radiation Physics, University of Gent, 9000 Gent, Belgium}}
\def\groupgiessen{\affiliation{Physikalisches Institut, Universit\"at Gie{\ss}en, 35392 Gie{\ss}en, Germany}}
\def\groupglasgow{\affiliation{SUPA, School of Physics and Astronomy, University of Glasgow, Glasgow G12 8QQ, United Kingdom}}
\def\groupillinois{\affiliation{Department of Physics, University of Illinois, Urbana, Illinois 61801-3080, USA}}
\def\groupmichigan{\affiliation{Randall Laboratory of Physics, University of Michigan, Ann Arbor, Michigan 48109-1040, USA }}
\def\groupmoscow{\affiliation{Lebedev Physical Institute, 117924 Moscow, Russia}}
\def\groupnikhef{\affiliation{National Institute for Subatomic Physics (Nikhef), 1009 DB Amsterdam, The Netherlands}}
\def\groupstpetersburg{\affiliation{Petersburg Nuclear Physics Institute, Gatchina, Leningrad region, 188300 Russia}}
\def\groupprotvino{\affiliation{Institute for High Energy Physics, Protvino, Moscow region, 142281 Russia}}
\def\groupregensburg{\affiliation{Institut f\"ur Theoretische Physik, Universit\"at Regensburg, 93040 Regensburg, Germany}}
\def\grouprome{\affiliation{Istituto Nazionale di Fisica Nucleare, Sezione Roma 1, Gruppo Sanit\`a and Physics Laboratory, Istituto Superiore di Sanit\`a, 00161 Roma, Italy}}
\def\grouptriumf{\affiliation{TRIUMF, Vancouver, British Columbia V6T 2A3, Canada}}
\def\grouptokyo{\affiliation{Department of Physics, Tokyo Institute of Technology, Tokyo 152, Japan}}
\def\groupamsterdam{\affiliation{Department of Physics and Astronomy, Vrije Universiteit, 1081 HV Amsterdam, The Netherlands}}
\def\groupwarsaw{\affiliation{Andrzej Soltan Institute for Nuclear Studies, 00-689 Warsaw, Poland}}
\def\groupyerevan{\affiliation{Yerevan Physics Institute, 375036 Yerevan, Armenia}}
\def\groupnone{\noaffiliation}

% Set Institute Order

\groupargonne
\groupbari
\groupbeijing
\groupcolorado
\groupdesy
\groupzeuthen
\groupdubna
\grouperlangen
\groupferrara
\groupfrascati
\groupgent
\groupgiessen
\groupglasgow
\groupillinois
\groupmichigan
\groupmoscow
\groupnikhef
\groupstpetersburg
\groupprotvino
\groupregensburg
\grouprome
\grouptriumf
\grouptokyo
\groupamsterdam
\groupwarsaw
\groupyerevan

% List of Authors

\author{A.~Airapetian}  \groupgiessen
\author{N.~Akopov}  \groupyerevan
\author{Z.~Akopov}  \groupdesy
\author{E.C.~Aschenauer}  \groupzeuthen
\author{W.~Augustyniak}  \groupwarsaw
\author{R.~Avakian}  \groupyerevan
\author{A.~Avetissian}  \groupyerevan
\author{E.~Avetisyan}  \groupdesy
\author{B.~Ball}  \groupmichigan
\author{S.~Belostotski}  \groupstpetersburg
\author{N.~Bianchi}  \groupfrascati
\author{H.P.~Blok}  \groupnikhef \groupamsterdam
\author{H.~B\"ottcher}  \groupzeuthen
\author{A.~Borissov}  \groupdesy
\author{J.~Bowles}  \groupglasgow
\author{I.~Brodski}  \groupgiessen
\author{V.~Bryzgalov}  \groupprotvino
\author{J.~Burns}  \groupglasgow
\author{M.~Capiluppi}  \groupferrara
\author{G.P.~Capitani}  \groupfrascati
\author{E.~Cisbani}  \grouprome
\author{G.~Ciullo}  \groupferrara
\author{M.~Contalbrigo}  \groupferrara
\author{P.F.~Dalpiaz}  \groupferrara
\author{W.~Deconinck}  \groupdesy
\author{R.~De~Leo}  \groupbari
\author{L.~De~Nardo}  \groupdesy \grouptriumf
\author{E.~De~Sanctis}  \groupfrascati
\author{M.~Diefenthaler}  \grouperlangen
\author{P.~Di~Nezza}  \groupfrascati
\author{J.~Dreschler}  \groupnikhef
\author{M.~D\"uren}  \groupgiessen
\author{M.~Ehrenfried}  \groupgiessen
\author{G.~Elbakian}  \groupyerevan
\author{F.~Ellinghaus}  \groupcolorado
\author{R.~Fabbri}  \groupzeuthen
\author{A.~Fantoni}  \groupfrascati
\author{L.~Felawka}  \grouptriumf
\author{S.~Frullani}  \grouprome
\author{D.~Gabbert}  \groupzeuthen
\author{G.~Gapienko}  \groupprotvino
\author{V.~Gapienko}  \groupprotvino
\author{F.~Garibaldi}  \grouprome
\author{G.~Gavrilov}  \groupdesy \groupstpetersburg \grouptriumf
\author{V.~Gharibyan}  \groupyerevan
\author{F.~Giordano}  \groupferrara
\author{S.~Gliske}  \groupmichigan
\author{M.~Golembiovskaya}  \groupdesy
\author{C.~Hadjidakis}  \groupfrascati
\author{M.~Hartig}  \groupdesy
\author{D.~Hasch}  \groupfrascati
\author{T.~Hasegawa}  \grouptokyo
\author{G.~Hill}  \groupglasgow
\author{A.~Hillenbrand}  \groupzeuthen
\author{M.~Hoek}  \groupglasgow
\author{Y.~Holler}  \groupdesy
\author{I.~Hristova}  \groupzeuthen
\author{Y.~Imazu}  \grouptokyo
\author{A.~Ivanilov}  \groupprotvino
\author{A.~Izotov}  \groupstpetersburg
\author{H.E.~Jackson}  \groupargonne
\author{A.~Jgoun}  \groupstpetersburg
\author{H.S.~Jo}  \groupgent
\author{S.~Joosten}  \groupillinois
\author{R.~Kaiser}  \groupglasgow
\author{G.~Karyan}  \groupyerevan
\author{T.~Keri}  \groupgiessen
\author{E.~Kinney}  \groupcolorado
\author{A.~Kisselev}  \groupstpetersburg
\author{N.~Kobayashi}  \grouptokyo
\author{V.~Korotkov}  \groupprotvino
\author{V.~Kozlov}  \groupmoscow
\author{P.~Kravchenko}  \groupstpetersburg
\author{V.G.~Krivokhijine}  \groupdubna
\author{L.~Lagamba}  \groupbari
\author{R.~Lamb}  \groupillinois
\author{L.~Lapik\'as}  \groupnikhef
\author{I.~Lehmann}  \groupglasgow
\author{P.~Lenisa}  \groupferrara
\author{L.A.~Linden-Levy}  \groupillinois
\author{A.~L\'opez~Ruiz}  \groupgent
\author{W.~Lorenzon}  \groupmichigan
\author{X.-G.~Lu}  \groupdesy
\author{B.-Q.~Ma}  \groupbeijing
\author{D.~Mahon}  \groupglasgow
\author{N.C.R.~Makins}  \groupillinois
\author{S.I.~Manaenkov}  \groupstpetersburg
\author{L.~Manfr\'e}  \grouprome
\author{Y.~Mao}  \groupbeijing
\author{B.~Marianski}  \groupwarsaw
\author{A.~Martinez de la Ossa}  \groupcolorado
\author{H.~Marukyan}  \groupyerevan
\author{C.A.~Miller}  \grouptriumf
\author{Y.~Miyachi}  \grouptokyo
\author{A.~Movsisyan}  \groupyerevan
\author{V.~Muccifora}  \groupfrascati
\author{M.~Murray}  \groupglasgow
\author{A.~Mussgiller}  \grouperlangen
\author{E.~Nappi}  \groupbari
\author{Y.~Naryshkin}  \groupstpetersburg
\author{A.~Nass}  \grouperlangen
\author{M.~Negodaev}  \groupzeuthen
\author{W.-D.~Nowak}  \groupzeuthen
\author{L.L.~Pappalardo}  \groupferrara
\author{R.~Perez-Benito}  \groupgiessen
\author{N.~Pickert}  \grouperlangen
\author{M.~Raithel}  \grouperlangen
\author{P.E.~Reimer}  \groupargonne
\author{A.R.~Reolon}  \groupfrascati
\author{C.~Riedl}  \groupzeuthen
\author{K.~Rith}  \grouperlangen
\author{G.~Rosner}  \groupglasgow
\author{A.~Rostomyan}  \groupdesy
\author{J.~Rubin}  \groupargonne
\author{D.~Ryckbosch}  \groupgent
\author{Y.~Salomatin}  \groupprotvino
\author{F.~Sanftl}  \groupregensburg
\author{A.~Sch\"afer}  \groupregensburg
\author{G.~Schnell}  \groupgent
\author{K.P.~Sch\"uler}  \groupdesy
\author{B.~Seitz}  \groupglasgow
\author{T.-A.~Shibata}  \grouptokyo
\author{V.~Shutov}  \groupdubna
\author{M.~Stancari}  \groupferrara
\author{M.~Statera}  \groupferrara
\author{E.~Steffens}  \grouperlangen
\author{J.J.M.~Steijger}  \groupnikhef
\author{H.~Stenzel}  \groupgiessen
\author{J.~Stewart}  \groupzeuthen
\author{F.~Stinzing}  \grouperlangen
\author{S.~Taroian}  \groupyerevan
\author{A.~Terkulov}  \groupmoscow
\author{A.~Trzcinski}  \groupwarsaw
\author{M.~Tytgat}  \groupgent
\author{A.~Vandenbroucke}  \groupgent
\author{P.B.~van~der~Nat}  \groupnikhef
\author{Y.~Van~Haarlem}  \groupgent
\author{C.~Van~Hulse}  \groupgent
\author{M.~Varanda}  \groupdesy
\author{D.~Veretennikov}  \groupstpetersburg
\author{V.~Vikhrov}  \groupstpetersburg
\author{I.~Vilardi}  \groupbari
\author{C.~Vogel}  \grouperlangen
\author{S.~Wang}  \groupbeijing
\author{S.~Yaschenko}  \grouperlangen
\author{H.~Ye}  \groupbeijing
\author{Z.~Ye}  \groupdesy
\author{S.~Yen}  \grouptriumf
\author{W.~Yu}  \groupgiessen
\author{V.~Zagrebelnyy}  \groupdesy \groupgiessen
\author{D.~Zeiler}  \grouperlangen
\author{B.~Zihlmann}  \groupdesy
\author{P.~Zupranski}  \groupwarsaw

\collaboration{The HERMES Collaboration} \noaffiliation

\date{\today}

\begin{abstract}
% insert abstract here
\end{abstract}

% insert suggested PACS numbers in braces
\pacs{}

% maketitle must be placed here
\maketitle

\end{document}