### 2011-12-12 frame_dual.tex
started going through the paper and changing to collaboration wishes

\documentclass[11pt,a4paper]{article}
\usepackage{jheppub}
\pdfoutput=1
\toccontinuoustrue
\usepackage{lscape}
\usepackage{vruler}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{appendix}

\newcommand{\blue}[1]{{\color{blue}{#1}\color{black}}}
\newcommand{\red}[1]{{\color{red}{#1}\color{black}}}

\def\desy{{\sc desy}}

\begin{document}
\title{Beam-Helicity and Beam-Charge Asymmetries \\
        in the Bethe--Heitler and DVCS processes on \\
        an Unpolarised Proton Target
  	\underline{Paper Tag: PDVCS0607 (v2.0)}
        \underline{Drafting Committee: DC90 (\today)}
}

% List of Institute Addresses 
\collaboration{The {\sc Hermes} Collaboration}
%\collaborationImg{\includegraphics{hermeslogo.pdf}}

\author[12]{A.~Airapetian}
\author[26]{N.~Akopov}
\author[5]{Z.~Akopov}
\author[6]{E.C.~Aschenauer}
\author[25]{W.~Augustyniak}
\author[26]{R.~Avakian}
\author[26]{A.~Avetissian}
\author[5]{E.~Avetisyan}
\author[15]{B.~Ball}
\author[18]{S.~Belostotski}
\author[10]{N.~Bianchi}
\author[17,24]{H.P.~Blok}
\author[6]{H.~B\"ottcher}
\author[5]{A.~Borissov}
\author[13]{J.~Bowles}
\author[12]{I.~Brodski}
\author[19]{V.~Bryzgalov}
\author[13]{J.~Burns}
\author[9]{M.~Capiluppi}
\author[10]{G.P.~Capitani}
\author[21]{E.~Cisbani}
\author[9]{G.~Ciullo}
\author[9]{M.~Contalbrigo}
\author[9]{P.F.~Dalpiaz}
\author[5]{W.~Deconinck}
\author[2]{R.~De~Leo}
\author[5,22,11]{L.~De~Nardo}
\author[10]{E.~De~Sanctis}
\author[8]{M.~Diefenthaler}
\author[10]{P.~Di~Nezza}
\author[17]{J.~Dreschler}
\author[12]{M.~D\"uren}
\author[12]{M.~Ehrenfried}
\author[26]{G.~Elbakian}
\author[4]{F.~Ellinghaus}
\author[6]{R.~Fabbri}
\author[10]{A.~Fantoni}
\author[22]{L.~Felawka}
\author[21]{S.~Frullani}
\author[6]{D.~Gabbert}
\author[19]{G.~Gapienko}
\author[19]{V.~Gapienko}
\author[21]{F.~Garibaldi}
\author[5,18,22]{G.~Gavrilov}
\author[26]{V.~Gharibyan}
\author[9]{F.~Giordano}
\author[15]{S.~Gliske}
\author[5]{M.~Golembiovskaya}
\author[10]{C.~Hadjidakis}
\author[5]{M.~Hartig}
\author[10]{D.~Hasch}
\author[23]{T.~Hasegawa}
\author[13]{G.~Hill}
\author[6]{A.~Hillenbrand}
\author[13]{M.~Hoek}
\author[5]{Y.~Holler}
\author[6]{I.~Hristova}
\author[23]{Y.~Imazu}
\author[19]{A.~Ivanilov}
\author[18]{A.~Izotov}
\author[1]{H.E.~Jackson}
\author[18]{A.~Jgoun}
\author[11]{H.S.~Jo}
\author[14]{S.~Joosten}
\author[13]{R.~Kaiser}
\author[26]{G.~Karyan}
\author[12]{T.~Keri}
\author[4]{E.~Kinney}
\author[18]{A.~Kisselev}
\author[23]{N.~Kobayashi}
\author[19]{V.~Korotkov}
\author[16]{V.~Kozlov}
\author[18]{P.~Kravchenko}
\author[7]{V.G.~Krivokhijine}
\author[2]{L.~Lagamba}
\author[14]{R.~Lamb}
\author[17]{L.~Lapik\'as}
\author[13]{I.~Lehmann}
\author[9]{P.~Lenisa}
\author[14]{L.A.~Linden-Levy}
\author[11]{A.~L\'opez~Ruiz}
\author[15]{W.~Lorenzon}
\author[5]{X.-G.~Lu}
\author[23]{X.-R.~Lu}
\author[3]{B.-Q.~Ma}
\author[13]{D.~Mahon}
\author[14]{N.C.R.~Makins}
\author[18]{S.I.~Manaenkov}
\author[21]{L.~Manfr\'e}
\author[3]{Y.~Mao}
\author[25]{B.~Marianski}
\author[4]{A.~Martinez de la Ossa}
\author[26]{H.~Marukyan}
\author[22]{C.A.~Miller}
\author[23]{Y.~Miyachi}
\author[26]{A.~Movsisyan}
\author[10]{V.~Muccifora}
\author[13]{M.~Murray}
\author[8]{A.~Mussgiller}
\author[2]{E.~Nappi}
\author[18]{Y.~Naryshkin}
\author[8]{A.~Nass}
\author[6]{M.~Negodaev}
\author[6]{W.-D.~Nowak}
\author[9]{L.L.~Pappalardo}
\author[12]{R.~Perez-Benito}
\author[8]{N.~Pickert}
\author[8]{M.~Raithel}
\author[1]{P.E.~Reimer}
\author[10]{A.R.~Reolon}
\author[6]{C.~Riedl}
\author[8]{K.~Rith}
\author[13]{G.~Rosner}
\author[5]{A.~Rostomyan}
\author[14]{J.~Rubin}
\author[11]{D.~Ryckbosch}
\author[19]{Y.~Salomatin}
\author[20]{F.~Sanftl}
\author[20]{A.~Sch\"afer}
\author[11]{G.~Schnell}
\author[5]{K.P.~Sch\"uler}
\author[13]{B.~Seitz}
\author[23]{T.-A.~Shibata}
\author[7]{V.~Shutov}
\author[9]{M.~Stancari}
\author[9]{M.~Statera}
\author[8]{E.~Steffens}
\author[17]{J.J.M.~Steijger}
\author[12]{H.~Stenzel}
\author[6]{J.~Stewart}
\author[8]{F.~Stinzing}
\author[26]{S.~Taroian}
\author[16]{A.~Terkulov}
\author[25]{A.~Trzcinski}
\author[11]{M.~Tytgat}
\author[11]{A.~Vandenbroucke}
\author[17]{P.B.~van~der~Nat}
\author[11]{Y.~Van~Haarlem}
\author[11]{C.~Van~Hulse}
\author[5]{M.~Varanda}
\author[18]{D.~Veretennikov}
\author[18]{V.~Vikhrov}
\author[2]{I.~Vilardi}
\author[8]{C.~Vogel}
\author[3]{S.~Wang}
\author[8]{S.~Yaschenko}
\author[3]{H.~Ye}
\author[5]{Z.~Ye}
\author[22]{S.~Yen}
\author[12]{W.~Yu}
\author[8]{D.~Zeiler}
\author[5]{B.~Zihlmann}
\author[25]{P.~Zupranski}

\affiliation[1]{Physics Division, Argonne National Laboratory, Argonne, Illinois 60439-4843, USA}
\affiliation[2]{Istituto Nazionale di Fisica Nucleare, Sezione di Bari, 70124 Bari, Italy}
\affiliation[3]{School of Physics, Peking University, Beijing 100871, China}
\affiliation[4]{Nuclear Physics Laboratory, University of Colorado, Boulder, Colorado 80309-0390, USA}
\affiliation[5]{DESY, 22603 Hamburg, Germany}
\affiliation[6]{DESY, 15738 Zeuthen, Germany}
\affiliation[7]{Joint Institute for Nuclear Research, 141980 Dubna, Russia}
\affiliation[8]{Physikalisches Institut, Universit\"at Erlangen-N\"urnberg, 91058 Erlangen, Germany}
\affiliation[9]{Istituto Nazionale di Fisica Nucleare, Sezione di Ferrara and Dipartimento di Fisica, Universit\`a di Ferrara, 44100 Ferrara, Italy}
\affiliation[10]{Istituto Nazionale di Fisica Nucleare, Laboratori Nazionali di Frascati, 00044 Frascati, Italy}
\affiliation[11]{Department of Subatomic and Radiation Physics, University of Gent, 9000 Gent, Belgium}
\affiliation[12]{Physikalisches Institut, Universit\"at Gie{\ss}en, 35392 Gie{\ss}en, Germany}
\affiliation[13]{SUPA, School of Physics and Astronomy, University of Glasgow, Glasgow G12 8QQ, United Kingdom}
\affiliation[14]{Department of Physics, University of Illinois, Urbana, Illinois 61801-3080, USA}
\affiliation[15]{Randall Laboratory of Physics, University of Michigan, Ann Arbor, Michigan 48109-1040, USA }
\affiliation[16]{Lebedev Physical Institute, 117924 Moscow, Russia}
\affiliation[17]{National Institute for Subatomic Physics (Nikhef), 1009 DB Amsterdam, The Netherlands}
\affiliation[18]{Petersburg Nuclear Physics Institute, Gatchina, Leningrad region, 188300 Russia}
\affiliation[19]{Institute for High Energy Physics, Protvino, Moscow region, 142281 Russia}
\affiliation[20]{Institut f\"ur Theoretische Physik, Universit\"at Regensburg, 93040 Regensburg, Germany}
\affiliation[21]{Istituto Nazionale di Fisica Nucleare, Sezione Roma 1, Gruppo Sanit\`a and Physics Laboratory, Istituto Superiore di Sanit\`a, 00161 Roma, Italy}
\affiliation[22]{TRIUMF, Vancouver, British Columbia V6T 2A3, Canada}
\affiliation[23]{Department of Physics, Tokyo Institute of Technology, Tokyo 152, Japan}
\affiliation[24]{Department of Physics and Astronomy, Vrije Universiteit, 1081 HV Amsterdam, The Netherlands}
\affiliation[25]{National Center for Nuclear Research, 00-689 Warsaw, Poland}
\affiliation[26]{Yerevan Physics Institute, 375036 Yerevan, Armenia}

\abstract{
Beam-helicity and beam-charge asymmetries in the hard exclusive leptoproduction of real photons from an unpolarised proton target by a 27.6\,GeV lepton beam 
are extracted from the H{\sc ermes} data set of 2006-2007 using a
missing-mass event selection technique. The beam-charge asymmetry
amplitudes extracted from the new data set are more
precise than those extracted from the earlier data
set of 1996-2005 previously analysed in the same manner by \red{\sc{Hermes}}. Results from the two data sets are found to be
compatible with each other. Combined results from these data sets are
extracted and represent the most statistically precise asymmetry amplitude measurements available in the H{\sc ermes} kinematic region.
}
\keywords{lepton-nucleon scattering}

\setvruler[15pt][1][1][3][1][30pt][30pt][-42pt]
\maketitle

\include{theory}
\include{experiment_and_data}
\include{background_and_systematic_dual}
\include{results_dual}
\include{summary}
\include{append_dual}
\bibliographystyle{natbib}
\bibliography{jburnsref}
\end{document}